---
title: "GIS-02 - More with Leaflet"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook extends our `leaflet` experience to include both multi-layer maps as well as thematic choropleth maps.

## Dependencies
This notebook requires a variety of packages for working with spatial data:

```{r load-packages}
# tidyverse packages
library(dplyr)        # data wrangling

# spatial packages
library(leaflet)      # interactive maps
library(mapview)      # preview spatial data
library(sf)           # spatial data tools
library(tigris)       # TIGER/Line data access

# other packages
library(here)         # file path management
library(RColorBrewer) # color palettes
```

## Load Data
First, we need to load up our data. This is review from the first session, and so we've pre-filled the code for you. We'll automatically re-project the data as well (also review):

```{r load-data}
# point data 1 - Violent Crime in the Shaw Neighborhood
shawCrime <- st_read(here("data", "SHAW_Violent_2018", "SHAW_Violent_2018.shp")) %>%
  st_transform(crs = 4326)

# point data 2 - Food Retail in St. Louis
grocery <- st_read(here("data", "STL_FOOD_Grocery.geojson")) 
cornerStore <- st_read(here("data", "STL_FOOD_Convenience.geojson"))

# polygon data 1 - St. Louis Neighborhood Population
nhood <- st_read(here("data", "STL_DEMOS_Nhoods", "STL_DEMOS_Nhoods.shp")) %>%
  st_transform(crs = 4326)

# polygon data 2 - St. Louis COVID Cases by ZIP Code
covid <- st_read(here("data", "daily_snapshot_regional.geojson"))
```

Now, take a few moments using your *console* and explore each of these data sets using the `mapview()` function!

## Download a Bit of Extra Data
We'll supplement the data we've loaded with some boundary data for a four county region in the St. Louis area (the City, the County, St. Charles County, and Jefferson County). 

We'll use the `tigris` package to do this. Using `tigris`, you can download most of the common spatial data sets published by the U.S. Census Bureau that reflect both administrative boundaries as well as elements of our physical and human geography. However, this package cannot be used to download population data (that is the purpose of `tidycensus`).

The pipelines read as follows:

1. We'll assign to our new object `region` the result of the following pipeline.
2. First, we'll download county boundaries for Missouri (FIPS Code 29), *then*
3. We'll select the `GEOID` and `NAMELSAD` columns using the `dplyr` `select()` function, *then*
4. We'll keep only the observations that match the GEOID codes for the select counties using the `dplyr` `filter()` function, *then*
5. We'll re-project our data.

```{r load-counties}
region <- counties(state = 29) %>%
  select(GEOID, NAMELSAD) %>%
  filter(GEOID %in% c("29099", "29183", "29189", "29510")) %>%
  st_transform(crs = 4326)

city <- filter(region, GEOID == "29510")
```

## More with Point Data
Last session, we made some simple point maps. Today, we want to extend those skills by mapping multiple point layers. We're going to start with the Shaw crime data and build a map together. First, we need to separate out two categories of crimes - homicides and aggravated assaults. We'll use the `dplyr` `filter()` function to do this:

```{r subset-shaw}
shawHomicide <- filter(shawCrime, crimeCt == "Homicide")
shawAssault <- filter(shawCrime, crimeCt == "Aggravated Assault")
```

Then, we'll use a number of different `leaflet` functions to build our map.

Within `addCircleMarkers`, we we'll use the following arguments:

  * `radius` - the size of the marker
  * `opacity` - the inner part of the marker
  * `color` - the color of the outer part of the marker
  * `fillColor` - the color of the inner part of the marker

```{r map-crime}

```

Now it is your turn, use the same basic process to map grocery stores and corner stores in St. Louis City and St. Louis County:

```{r map-stores}

```

## Making Choropleth Maps
We can also create thematic choropleth maps that map quantities using `leaflet`. We'll use this as a way to demonstrate how to overlay polygons on top of each other as well.

Instead of `addCircleMarkers()` or `addMarkers()`, we'll use `addPolygons()` with some of the following options:

* `color` - outline ("stroke") color for each polygon
* `weight` - stroke width
* `opacity` - stroke opacity
* `smoothFactor` - allows `leaflet` to simplify polygons depending on zoom
* `fillOpacity` - fill opacity
* `fillColor` - creates the fill itself
* `highlightOptions` - creates effect when mouse drags over specific polygons

```{r map-population}

```

Now it is your turn. Apply the same principles to mapping COVID cases in St. Louis:

```{r map-covid}

```

